---
title: "heart_rate_analysis"
author: "John Goldin"
date: "2/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r setup_libs, echo = FALSE}
library(tidyverse)
library(jsonlite)
library(lubridate)
library(ggplot2)
library(scales)
library(XML)
```

```{r load_data, cache = TRUE, echo = FALSE, eval = FALSE}
# load("full_data_from_Apple_Health.RData")
# Using example from github: https://gist.github.com/ryanpraski/ba9baee2583cfb1af88ca4ec62311a3d
# ryanpraski/apple_health_load_analysis_R.r
# http://www.ryanpraski.com/apple-health-data-how-to-export-analyze-visualize-guide/

# see also:  https://github.com/deepankardatta/AppleHealthAnalysis

#load apple health export.xml file
xml <- xmlParse("apple_health_export/export.xml")

#json_example <- read_json("apple_health_export/export.xml")
#transform xml file to data frame - select the Record rows from the xml file
df <- XML:::xmlAttrsToDataFrame(xml["//Record"])

#make value variable numeric
df$value <- as.numeric(as.character(df$value))


#make endDate in a date time variable POSIXct using lubridate with eastern time zone
df$endDate <-ymd_hms(df$endDate,tz="America/New_York")
df$startDate <-ymd_hms(df$startDate,tz="America/New_York")

##add in year month date dayofweek hour columns
df$month<-format(df$endDate,"%m")
df$year<-format(df$endDate,"%Y")
df$date<-format(df$endDate,"%Y-%m-%d")
df$dayofweek <-wday(df$endDate, label=TRUE, abbr=FALSE)
df$hour <-format(df$endDate,"%H")

steps <- df %>%
  filter(type == 'HKQuantityTypeIdentifierStepCount') %>%
  group_by(date) %>% 
  mutate(steps = as.numeric(as.character(value))) %>%
  summarise(steps = sum(steps))
distance <- df %>%
  filter(type == 'HKQuantityTypeIdentifierDistanceWalkingRunning') %>%
  group_by(date) %>% 
  mutate(distance = as.numeric(as.character(value))) %>%
  summarise(distance = sum(distance))
#filter(steps, month(date) == 5, year(date) == 2016) %>% print(n=1000)

```
```{r create_plot_function, echo = FALSE}
plot_day <- function(adf, adate, events = FALSE, hour_min = NULL, hour_max = NULL) {
  browser()
  print(hour_max)
  adf <- adf %>% filter(type %in% c("HKQuantityTypeIdentifierHeartRate", "HeartRate"), local_date == adate) %>%
    arrange(local_start)
  if (!is.null(hour_max)) adf <- adf %>% filter(hour(local_start) <= hour_max)
  if (!is.null(hour_min)) adf <- adf %>% filter(hour(local_start) >= hour_min)
  p <- ggplot(data = adf, aes(x = local_start, y = value))
  if (events) {
    base <- adf$local_start[10]
    left_house <- base
    hour(left_house) <- 5
    minute(left_house) <- 55
    arrived <- base
    hour(arrived) <- 6
    minute(arrived) <- 3
    p <- p + geom_vline(xintercept = left_house, colour = "yellow") +
      geom_vline(xintercept = arrived, colour = "yellow")
  }
  p <- p + geom_point(size = 0.3) +
    xlab("Time (Hour)") + ylab("Pulse (Apple Watch)") +
    scale_x_datetime(date_labels = "%H", breaks = date_breaks("1 hour"),
                     minor_breaks = date_breaks("10 min")) + 
    facet_wrap(~ local_date, ncol = 1, scales = "free_x")
}
```


https://www.cnn.com/2019/03/16/health/apple-watch-heart-rate-arrhythmia-afib-study/index.html
https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6111985/

```{r recreate_oct_2020_plot}
load("~/Dropbox/Programming/R_Stuff/john_vitals/Apple-Health-Data/key_days_export.RData")
p <- plot_day(day_2020_10_25_health, c(ymd("2020-10-25")), events = TRUE, hour_min = 8, hour_max = 15)

days_events <- tribble(
  ~x, ~lab, 
  as_datetime("2020-10-25 12:30:00"), "ECG",
  as_datetime("2020-10-25 13:47:00"), "ECG"
)
ynhh <- tribble(
  ~x, ~lab,
  as_datetime("2020-10-25 14:04:00"),
  "arrive at ER"
)
p <- p  + geom_segment(data = days_events, aes(x = x, xend = x, y = 154, yend = 159.3), colour = "orange", arrow = arrow(angle = 20, ends = "first", length = unit(0.03, "npc"))) + 
  geom_text(data = days_events, aes(x = x, y = 155, label = lab), hjust = 0.5, color = "orange", nudge_y = 5, size = 2) +
    geom_segment(data = ynhh, aes(x = x, xend = x, y = 154, yend = 159.3), colour = "blue", arrow = arrow(angle = 20, ends = "first", length = unit(0.03, "npc"))) +
  geom_text(data = ynhh, aes(x = x, y = 155, label = lab), hjust = 0, color = "blue", nudge_y = 5, size = 3) +
  annotate("label", y = 130, x = as_datetime("2020-10-25 11:11:00"),
           xmin =  as_datetime("2020-10-25 10:41:00"), xmax = as_datetime("2020-10-25 11:50:00"),
           label = "hike in the woods", size = 3)
ggsave("pdfs/ER visit John Goldin 10-25-2020-alt.pdf", width = 6.5, height = 4.5 )

ptoday <- plot_day(day_2020_10_25_health, adate = as_date("2020-10-25"), events = TRUE, hour_min = 8, hour_max = 15)


# health_df %>% filter(as_date(start_date) ==  ymd("2019-02-16")) %>% arrange(start_date) %>% View()
```



```{r}
plot_day(health_df, ymd("2019-02-15"), events = TRUE, hour_min = 1, hour_max = 2) %>% print()

# health_df %>% filter(as_date(start_date) ==  ymd("2019-02-16")) %>% arrange(start_date) %>% View()
```


```{r}
plot_day(health_df, ymd("2020-10-25"),  hour_min = 8, hour_max = 15) %>% print()
plot_day(health_df, ymd("2020-11-01"),  hour_min = 6, hour_max = 12) %>% print()


# health_df %>% filter(as_date(start_date) ==  ymd("2019-02-16")) %>% arrange(start_date) %>% View()
```



```{r anomalous_readings, echo = FALSE, eval = FALSE}
df <- df  %>% filter(type == "HKQuantityTypeIdentifierHeartRate", start_date >= ymd("2017-10-01")) 
above145 <- df %>% filter(type == "HKQuantityTypeIdentifierHeartRate", value >= 145 ) %>%
  arrange(start_date) %>%
  select(start_date, value, dayofweek, hour, date, type, year)
table(above145$date)
plot_day(df, "2019-02-02") %>% print()

above145_night <- df %>% filter(type == "HKQuantityTypeIdentifierHeartRate", value >= 145,
                                (as.numeric(hour) <= 6) | (as.numeric(hour) >= 23)) %>%
                                # (hour <= 6) ) %>%
  arrange(start_date) %>%
  select(start_date, value, dayofweek, hour, date, type, year)
# This gives us four dates, one of which is 2/15/2019.
```


```{r summary}
p15th <- plot_day(df, "2019-02-15", hour_min = 0, hour_max = 24)

weird_dates <- c("2018-08-24", "2018-08-26", "2018-03-27", "2018-06-13")
weird_plot <- plot_day(df, weird_dates, hour_min = 0, hour_max = 24)
strenuous <- plot_day(df, c("2018-03-24", "2018-06-12", "2018-04-13", "2018-04-08"), hour_min = 0, hour_max = 24)
```


```{r}
vhr <- df %>% filter(type == "HKQuantityTypeIdentifierHeartRateVariabilitySDNN")
```
Where xx is heart rate only:
xx %>% filter(value >= 160) %>% select(major_version, startDate, local_start, value, workoutActivityType) %>% View()

Two readings of 162 on 3/13/2021. 19:27:06 and 19:27:08
two EKGs on 3/13 at 7:26 and 7:28, heart rate 74 and 63
Looking at the plot_day graph, frequent readings from 16:00:00 onward
and the 162 readings are total outliers (from abouit 78).
From EKG looks like I was doing EKG about that time so may have been
physically fiddling with the watch.

plot_day(health_df, ymd("2021-03-13")) %>% print()
plot_day(health_df, ymd("2020-06-21")) %>% print()
plot_day(health_df, ymd("2020-04-15")) %>% print()
plot_day(health_df, ymd("2020-04-17")) %>% print()
plot_day(health_df, ymd("2020-03-26")) %>% print()
plot_day(health_df, ymd("2019-04-01")) %>% print()
plot_day(health_df, ymd("2018-12-18"), hour_min = 11, hour_max = 17) %>% print()
plot_day(health_df, ymd("2018-12-16"), hour_min = 11, hour_max = 17) %>% print()
plot_day(health_df, ymd("2018-12-18"), hour_min = 11, hour_max = 17) %>% print()
plot_day(health_df, ymd("2020-10-25"), hour_min = 11, hour_max = 17) %>% print()
plot_day(health_df, ymd("2020-11-14")) %>% print() # 172

plot_day(health_df, ymd("2018-04-12")) %>% print()


Can we look for outliers? 

Look only at rows that are not in Worout. Remember that
for data earlier than 90 days there is not much 
opportunity to be an outlier. For recent data,
workouts swamp the count.

```{r}
hr <- health_df %>% 
  filter(str_detect(sourceName, "Watch"), type == "HeartRate", is.na(workoutActivityType))
outliers <- hr %>% 
  mutate(prior = lag(value) / value,
    outlier_type = case_when(
      (lag(value) < 0.6*value) & (lead(value) < 0.6*value) ~ "high",
      (lead(value) > 1.4*value) | (lag(value) > 1.4*value) ~ "low",
      TRUE ~ "none"))
out_table_low <- outliers %>% 
  filter(outlier_type != "high") %>% 
  tabyl(outlier_type, major_version)
out_table_low %>% chisq.test(tabyl_results = TRUE)




View(outliers %>% select(local_start, local_end, prior, value,  sourceVersion, timezone, major_version))

xx <- hr %>% 
  mutate(outlier = case_when(
    value < 40 ~ "<= 39",
    value >= 160 ~ ">= 160",
    value >= 150 ~ "150-159",
    TRUE ~ "common"
  ))

xx %>% filter(value >= 160) %>% select(major_version, local_start, local_end, value, workoutActivityType) %>% View()

hr %>% filter(value > 150, major_version == "7") %>% View()

hr %>% filter(value > 120, Period == "Sleep") %>% View()
```

