---
title: "Notes on Intra-day Apple Health Export"
output: html_notebook
---

Outside Online article on watch:
https://www.outsideonline.com/2006026/apples-fitness-guru-opens-about-watch#close
(with ABC video of Apple lab)

Apple Watch 6 health and fitness press release
https://www.apple.com/newsroom/2019/06/watchos-6-advances-health-and-fitness-capabilities-for-apple-watch/

forensic detail about health kit:  https://blog.elcomsoft.com/2019/01/securing-and-extracting-health-data-apple-health-vs-google-fit/

See [Harris-Bennett formula for basal energy](https://en.wikipedia.org/wiki/Harris–Benedict_equation)

The Harris–Benedict equations revised by [Mifflin and St Jeor in 1990](https://pubmed.ncbi.nlm.nih.gov/2305711-a-new-predictive-equation-for-resting-energy-expenditure-in-healthy-individuals/):[4]

Men	BMR = (10 × weight in kg) + (6.25 × height in cm) - (5 × age in years) + 5
Women	BMR = (10 × weight in kg) + (6.25 × height in cm) - (5 × age in years) - 161

for me that would be 10 * pounds * 0.4536 + ( 6.25 * 69 * 2.54) - (5 * 70) + 5

```{r}
bmr <- (10 * 146 * 0.4536) + ( 6.25 * 69 * 2.54) - (5 * 69) + 5
print(bmr)
print(bmr * 1.76)

bmr / (24 * 60 * 60) # calories per second
```

basal energy per day by version

```{r}
intra %>% 
  filter(type == "Basal_Energy") %>% 
  group_by(sourceVersion) %>% 
  summarise(start = min(local_start, na.rm = TRUE), end = max(local_start, na.rm = TRUE),
            days = length(unique(date)), total_basal = sum(value, na.rm = TRUE), mean_basal = total_basal / days)
```

Some notes on events:
walk on 2/19/2020 15:02 to 16:04, no workout
2/23/2020 14:15 to 15:20 no workout
2/21/2020 no workout 13:15 to 14:20
no workout 2/28/2020 13:31 to 14:38

2/25/2020 removed watch 14:12 to 15:50
# example of long off period: utc 2019-09-18 21:18:55
2/27/2020 removed watch 08:56 to 09:16

battery ran out of power on 2/28/2020

version 5.0 start 2018-09-22 16:38:35
version 5.01 2018-09-30 13:34:32
version 5.1.1  2018-11-18 22:06:21

```{r}
no_workout <- intra4 %>% 
  filter(local_date %in% as_date(c("2020-02-19", "2020-02-21", "2020-02-23", "2020-02-28")),
         ((local_start >= as_datetime("2020-02-19 15:02:00")) & (local_start <= as_datetime("2020-02-19 16:04:00"))) |
         ((local_start >= as_datetime("2020-02-21 13:15:00"))  & (local_start <= as_datetime("2020-02-21 14:20:00"))) |
         ((local_start >= as_datetime("2020-02-28 13:31:00"))  & (local_start <= as_datetime("2020-02-28 14:38:00"))) |
         ((local_start >= as_datetime("2020-02-23 14:15:00"))  & (local_start <= as_datetime("2020-02-23 15:20:00"))))
feb_workouts <- workout_df %>% filter(year(local_start) == 2020, month(local_start) == 2)
yes_workout <- intra4 %>% 
  filter(local_date %in% as_date(c("2020-02-14", "2020-02-16", "2020-02-05", "2020-02-09")),
         ((local_start >= as_datetime("2020-02-14 09:55:29")) & (local_start <= as_datetime("2020-02-14 11:03:06"))) |
         ((local_start >= as_datetime("2020-02-16 10:40:22"))  & (local_start <= as_datetime("2020-02-16 11:51:46"))) |
         ((local_start >= as_datetime("2020-02-05 14:23:00"))  & (local_start <= as_datetime("2020-02-05 15:30:00"))) |
         ((local_start >= as_datetime("2020-02-09 14:18:00"))  & (local_start <= as_datetime("2020-02-09 15:28:00"))))
workout_sit <- intra4 %>% 
  filter(local_date %in% as_date(c("2020-02-29")),
         ((local_start >= as_datetime("2020-02-29 07:51:00")) & (local_start <= as_datetime("2020-02-29 08:58:00"))))
require(hms)
in_office <- intra4 %>% 
  filter(local_date%in% as_date(c("2020-02-19", "2020-02-24", "2020-02-10", "2020-02-12")),
        start_time >= as_hms("10:40:00"), start_time <= as_hms("11:50:00"))
yes_workout$workout <- "Workout"
no_workout$workout <- "Not Workout"
in_office$workout <- "In Office"
workout_sit$workout <- "Workout Sit"
work <- bind_rows(yes_workout, no_workout, in_office, workout_sit) %>% 
  filter(type2 != "Exercise_Watch") %>% 
  mutate(type2 = factor(type2, levels = c("Active_Energy", "Basal_Energy", "Heart_Rate",
                                        "Steps_Watch", "Walking_Watch", "Climb_Watch")),
         activity = case_when(workout %in% c("Workout", "Not Workout") ~ "Walking",
                       workout %in% c("Workout Sit", "In Office") ~ "Sitting"),
         workout = factor(workout, levels = c("Workout", "Workout Sit", "Not Workout", "In Office"),
                          labels = c("Walking\nWorkout", "Sitting\nWorkout", "Walk, No\nWorkout", "Sitting No\nWorkout")))


raw_table <- work %>% 
  group_by(type2, workout, activity, local_date) %>% 
  summarise(total = sum(value), n = n(), interval = median(interval, na.rm = TRUE), span = median(span, na.rm = TRUE)) %>% 
  mutate(total = case_when(
    type2 == "Heart_Rate" ~ total/n,
    TRUE ~ total
  ))

for_plot <- raw_table %>% group_by(type2, workout, activity) %>% 
  summarise(amount = mean(total), obs = mean(n), interval =mean(interval, na.rm = TRUE), span = mean(span, na.rm = TRUE)) %>% 
  mutate(amount = round(amount), obs = round(obs))

#https://stackoverflow.com/questions/30179442/plotting-minor-breaks-on-a-log-scale-with-ggplot/33179099#33179099
log10_minor_break = function (n_intervals = 9, ...){
  function(x) {
    minx         = floor(min(log10(x), na.rm=T))-1;
    maxx         = ceiling(max(log10(x), na.rm=T))+1;
    n_major      = maxx-minx+1;
    major_breaks = seq(minx, maxx, by=1)
    minor_breaks = 
      rep(log10(seq(1, n_intervals, by=1)), times = n_major)+
      rep(major_breaks, each = n_intervals)
    return(10^(minor_breaks))
  }
}
pmedian <- ggplot(data = for_plot %>% filter(type2 != "Climb_Watch", type2 != "Steps_Watch"), aes(y = interval, x =fct_rev(workout) )) + 
  facet_wrap(~ type2) + geom_col(aes(fill = fct_rev(activity))) +
  scale_y_log10(breaks = c(6, 60, 600),minor_breaks = log10_minor_break(6)) +
  # scale_y_log10(breaks = c(5, 60, 6000),minor_breaks = minor_breaks_n(5)) + 
  ylab("median interval between observations (log scale, in seconds)") + xlab('"workout" and actitivity')+
  theme(legend.position="top") +
  guides(fill=guide_legend(title="Activity: ")) +
  geom_text(aes(fill = NULL, label = round(interval)), hjust = 1) + 
  coord_flip() +
  ggtitle("How Does Interval Between Observations\nRelate to Declared Workout and Amount of Activity")
```
```{r display_workout_table, echo = FALSE}
print(pmedian)
```

[Apple Developer Guidelines](https://developer.apple.com/design/human-interface-guidelines/watchos/user-interaction/accelerometer-and-gyroscope/)
Use motion data to offer a tangible benefit to the user. 
The gathering of motion data should offer a tangible benefit to the user.
For example, fitness apps use the data to provide feedback about the user’s activity and general health.
Do not record data simply to have the data.

guy became suspect because of fitness tracking
https://www.nbcnews.com/news/us-news/google-tracked-his-bike-ride-past-burglarized-home-made-him-n1151761

Let's look at workout and basal energy. Do totals agree.
```{r total_by_workout}
y2020 <- intra4 %>% 
  filter(year(utc_start) == 2020, !is.na(workout_utc_start)) %>% 
  group_by(workout_utc_start, workout_utc_end) %>% 
  summarise(basal_sum = sum(value * (type == "BasalEnergyBurned"), na.rm = TRUE),
            active_sum = sum(value * (type == "ActiveEnergyBurned"), na.rm = TRUE),
            duration = last(duration)) %>% 
  left_join(workout_df %>% select(utc_start, utc_end, duration, totalEnergyBurned),
            by = c("workout_utc_start" = "utc_start", "workout_utc_end" = "utc_end"))
```

let's do a plot over a one week period of active and basal energy

```{r}
aday1 <- intra4 %>%
  filter(local_date == ymd("2020-02-09"))
aday <- aday1 %>% filter( hour >= 6, hour <= 21, hour >= 12, hour <= 16,
         type %in% c("ActiveEnergyBurned", "BasalEnergyBurned")) %>%
  mutate(end_time = as.integer(difftime(local_end, floor_date(local_end, "day"), unit = "secs")) %>% hms::hms()) %>%
  filter(end_time >= start_time) %>% arrange(type, start_time)
ggplot(data = aday,
       aes(x = start_time, y = intensity_raw, colour = type))  +
  # geom_point(data = aday %>% filter(type == "HeartRate"), size = 0.5) +
  geom_segment(data = aday %>% filter(type != "HeartRate"),
               aes(xend = end_time, yend = intensity_raw)) + ylim(0, 2)

pintensity <- ggplot(data = aday,
       aes(x = start_time, y = intensity, colour = type))  +
  # geom_point(data = aday %>% filter(type == "HeartRate"), size = 0.5) +
  geom_segment(data = aday %>% filter(type != "HeartRate"),
               aes(xend = end_time, yend = intensity, size = span)) + ylim(0, 2)

View(aday %>% filter(hour == 16) %>% select(local_start, span, value, type, interval, intensity, intensity_raw, workoutActivityType))

View(aday %>% filter(hour == 12) %>% select(local_start, span, value, type, interval, intensity, intensity_raw, workoutActivityType))
```

More detail on intra-day
```{r}
aday <- intra4 %>% 
  filter(local_date == ymd("2020-02-09"))
```

