---
title: "Summary of 2-15-2019 Event"
author: "John Goldin"
date: "2/19/2019"
output: 
  pdf_document:
    fig_height: 4.5
    fig_width: 6.5
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Source

Heart rate data is from the Apple Watch. I have readings from 2/20/2017, almost two years, and so far 
there are
almost 300,000 observations. The watch seems to take a pulse reading at least every five or six minutes.
During a workout (which I start on the watch), the frequency is much higher. It's also much more frequent
while I am looking at the heart display on the watch (as much as one reading every five seconds or so). 
It seems to have other ways to judge when more frequnt readings are desireable. 

The Apple support information describing how it measures heart rate is here: https://support.apple.com/en-us/HT204666

One can export all the data from the Apple Health app. The export is done as an XML file which
I am able to read into R via the XML package and analyze with standard tools.

```{r import_data, echo = TRUE, eval = FALSE, message = FALSE, warning = FALSE}
xml <- xmlParse("apple_health_export/export.xml")
df <- XML:::xmlAttrsToDataFrame(xml["//Record"])
hr <- df %>% filter(type == "HKQuantityTypeIdentifierHeartRate")
```

I use an Omron blood pressure cuff that has an app that moves blood pressure readings into
the Apple Health app. The same XML export has both heart rate and blood pressure as well as
other items such as step count, stairs climbed, exercise time, distsance walked/run, walking
heart rate average, heart rate variability, and resting heart rate.

The next plot shows the deatail of heart rate readings on the day I went to the ER. The first plot shows the period from
midnight to 11AM. The second plot shows detail from 5AM to 7AM. The vertical green lines mark the travel
period from our house to the Guilford ER. (The times are based on the Google Maps Timeline which shows to the minute when we
left and when we arrived. This chart represents some serious nerdiness!) The time interval between heart readings
varies. At about 1:30 I had a brief period of disgestive distress, which I won't describe in detail here. I went back to 
bed and my pulse settled down quickly. There was a gradual rise from about 1:45 until 5:40 when my pulse jumped from about 80 to about 108. This may have been when I got up to pee. I went back to bed and very shortly before 5:50 my pulse started to climb rapidly up to 150. The intervals between readings are short. Was I looking at my watch then or did the watch algorithm decide to 
take more readings because they were changing? I'm not sure. I know I did look at my watch and I saw the reading of 150.

In the bottom detail plot one can see that by the time we got into the car to drive to the ER my pulse was already decreasing.
Halfway through the car ride I did remark out loud that I was feeling a decrease in hearet rate. By the time we got into the 
ER my pulse was already on it's way back to normal. I would guess it was between 100 and 110 by the time they took my vitals,
but I don't know. The peak is very short duration. When I got to the ER I told them that I thought it had stayed at its high point
for 5 or 10 minutes. My perception and recollection was incorrect.

Note that there are two very anomalous data points, one at about 5:55 at about 108 BPM and another 
at 6:04 at 75 BPM. Are these just measurement errors or do they relate to something real? I don't know. I certainly have
observed cases where the Apple Watch produces a wacky reading, possibly because the watch is being jostled.

To get a better sense of the quality of the watch heart rate data, I looked at all days where there was a reading over 145, 54 out of 503 days of watch data. (There are a total of 296,000 heart rate readings in the dataset.) The next charts show some of those days with readings over 145. Some show a high peak in the midst of a period of heavy exercise.
(Via Google Maps Timeline sometimes I can even identify where I was hiking at the time of the reading.)
Some are just odd points that seem to come out of the blue. I did a separate search to look for days with a reading
over 145 when I would have been in bed. I only found three. One is the ER visit and the other two are the last two plots on the last page.
Now I realize that the one on April 25 was while I was hiking in Greece and the plot doesn't reflect the six or seven hour time difference.
The event on February 15 really was one of a kind even though the duration was very short.
\pagebreak

```{r setup_libs, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(jsonlite)
library(lubridate)
library(ggplot2)
library(XML)
library(scales)
# library(ggpubr)
```


```{r mods, echo = FALSE, message = FALSE, cache = TRUE, eval = FALSE}
xml <- xmlParse("apple_health_export/export.xml")
df <- XML:::xmlAttrsToDataFrame(xml["//Record"])
hr <- df %>% filter(type == "HKQuantityTypeIdentifierHeartRate")
#make value variable numeric
df$value <- as.numeric(as.character(df$value))


#make endDate in a date time variable POSIXct using lubridate with eastern time zone
df$endDate <-ymd_hms(df$endDate,tz="America/New_York")
df$startDate <-ymd_hms(df$startDate,tz="America/New_York")

##add in year month date dayofweek hour columns
df$month<-format(df$endDate,"%m")
df$year<-format(df$endDate,"%Y")
df$date<-format(df$endDate,"%Y-%m-%d")
df$dayofweek <-wday(df$endDate, label=TRUE, abbr=FALSE)
df$hour <-format(df$endDate,"%H")
```
```{r obs_per_month, echo = FALSE, eval = FALSE}
hr <- df %>% 
  filter(type == "HKQuantityTypeIdentifierHeartRate", startDate >= ymd("2017-10-01")) %>% 
  mutate(ym = str_sub(date, 1, 7)) 
p <- ggplot(data = hr, aes(x = ym)) + geom_bar()
p <- p + ylab("Number of Observations") + xlab("Month")
```

```{r create_plot, echo = FALSE}
plot_day <- function(adf, adate, events = FALSE, hour_min = NULL, hour_max = NULL, pt_size = 0.3) {
  browse()
  adf <- adf %>%
    filter(type == "HKQuantityTypeIdentifierHeartRate", date %in% adate) %>%
    arrange(startDate)
  if (!is.null(hour_max)) adf <- adf %>% filter(hour(startDate) <= hour_max)
  if (!is.null(hour_min)) adf <- adf %>% filter(hour(startDate) >= hour_min)
  p <- ggplot(data = adf, aes(x = startDate, y = value))
  if (events) {
    base <- adf$startDate[10]
    left_house <- base
    hour(left_house) <- 5
    minute(left_house) <- 55
    arrived <- base
    hour(arrived) <- 6
    minute(arrived) <- 3
    p <- p + geom_vline(xintercept = left_house, colour = "green") +
      geom_vline(xintercept = arrived, colour = "green")
  }
  p <- p + geom_point(size = pt_size) +
    xlab("Time (Hour)") + ylab("Pulse (Apple Watch)") +
    scale_x_datetime(date_labels = "%H", breaks = date_breaks("1 hour"),
                     minor_breaks = date_breaks("10 min")) + 
    facet_wrap(~ date, ncol = 1, scales = "free_x")
}
```

```{r plot1, echo = FALSE}

the_day <- plot_day(health_df_2019_02_15, "2019-02-15", events = TRUE, hour_max = 10)
the_day_detail <- plot_day(health_df, "2019-02-15", events = TRUE, hour_min = 5, hour_max = 6, pt_size = 0.7)
# the_day2 <- ggarrange(the_day, the_day_detail, ncol = 1)
print(the_day)
```

```{r plot2, echo = FALSE}
print(the_day_detail)
```

```{r anomalies, echo = FALSE, fig.height = 8.5}
weird_dates <- c("2018-08-24", "2018-08-26", "2018-03-27", "2019-02-02")
weird_plot <- plot_day(df, weird_dates, hour_min = 7, hour_max = 18)
strenuous <- plot_day(df, c("2018-03-24", "2018-06-12", "2018-04-13", "2018-04-08"), hour_min = 7, hour_max = 18)
night_weird <- plot_day(df, c("2017-10-16", "2018-08-09"), hour_min = 0, hour_max = 24)
other_weird <- plot_day(df, c("2018-04-25", "2018-10-04"), hour_min = 7, hour_max = 18)
print(weird_plot)
```
\pagebreak
```{r strenuous, echo = FALSE, fig.height = 8.5}
print(strenuous)
```

```{r plot3, echo = FALSE, fig_height = 4}
print(night_weird)
```

```{r plot4, echo = FALSE}
print(other_weird)
```
